---
title: 'CMTH 642 Data Analytics: Advanced Methods'
output:
  pdf_document: default
  html_document: default
  word_document: default
---
<center> <h1> Assignment 1 (10%) </h1> </center>
<center>  <h3> Rui Zhang </h2> </center>
<center> <h3> 500736315 </h2> </center>
---

#### 1. Read the csv files in the folder. (3 points)
```{r}
USDA_Macronutrients = read.csv("USDA_Macronutrients.csv")
USDA_Micronutrients = read.csv("USDA_Micronutrients.csv")
USDA_Macronutrients
USDA_Micronutrients
```

#### 2. Merge the data frames using the variable "ID". Name the Merged Data Frame "USDA". (6 points)
```{r}
USDA <- merge(USDA_Macronutrients, USDA_Micronutrients, by="ID")
```

#### 3. Check the datatypes of the attributes. Delete the commas in the Sodium and Potasium records. Assign Sodium and Potasium as numeric data types. (6 points)
```{r}
sapply(USDA, typeof)
USDA$Sodium<-gsub(",", "", USDA$Sodium)
USDA$Potassium<-gsub(",", "", USDA$Potassium)
USDA$Sodium<-as.numeric(USDA$Sodium)
USDA$Potassium<-as.numeric(USDA$Potassium)
#str(USDA)
```

#### 4. Remove records (rows) with missing values in more than 4 attributes (columns). How many records remain in the data frame? (6 points)
```{r}
USDA<-USDA[rowSums(is.na(USDA))<5, ]
```

#### 5. For records with missing values for Sugar, Vitamin E and Vitamin D, replace missing values with mean value for the respective variable. (6 points)
```{r}
mean.sugar<-mean(USDA$Sugar, na.rm = T)
mean.vitaminD<-mean(USDA$VitaminD, na.rm = T)
mean.vitaminE<-mean(USDA$VitaminE, na.rm = T)
USDA$Sugar[is.na(USDA$Sugar)]<-mean.sugar
USDA$VitaminD[is.na(USDA$VitaminD)]<-mean.vitaminD
USDA$VitaminE[is.na(USDA$VitaminE)]<-mean.vitaminE
```

#### 6. With a single line of code, remove all remaining records with missing values. Name the new Data Frame "USDAclean". How many records remain in the data frame? (6 points)
```{r}
USDAclean<-na.omit(USDA)
```

#### 7. Which food has the highest sodium level? (6 points)
```{r}
USDAclean[USDAclean$Sodium==max(USDAclean$Sodium),]$Description
```

#### 8. Create a histogram of Vitamin C distribution in foods. (6 points)
```{r}
hist(USDAclean$VitaminC, xlim=c(0, 300), ylim=c(0, 0.06), breaks=100, freq = F)
```

#### 9. Create a boxplot to illustrate the distribution of values for TotalFat, Protein and Carbohydrate. (6 points)
```{r}
nms<-c("TotalFat","Protein","Protein")
boxplot(USDAclean[nms], names=nms, col = c("orange","red", "yellow"))
```

#### 10. Create a scatterplot to illustrate the relationship between a food's TotalFat content and its Calorie content. (6 points)
```{r}
plot(USDAclean$TotalFat, USDAclean$Calories, col = "red")
```

#### 11. Add a variable to the data frame that takes value 1 if the food has higher sodium than average, 0 otherwise. Call this variable HighSodium. Do the same for High Calories, High Protein, High Sugar, and High Fat. How many foods have both high sodium and high fat? (8 points)
```{r}
mean.sodium<-mean(USDAclean$Sodium)
USDAclean$HighSodium<-ifelse(USDAclean$Sodium > mean.sodium, 1, 0)
mean.calories<-mean(USDAclean$Calories)
USDAclean$HighCalories<-ifelse(USDAclean$Calories > mean.calories, 1, 0)
mean.protein<-mean(USDAclean$Protein)
USDAclean$HighProtein<-ifelse(USDAclean$Protein > mean.protein, 1, 0)
mean.sugar<-mean(USDAclean$Sugar)
USDAclean$HighSugar<-ifelse(USDAclean$Sugar > mean.sugar, 1, 0)
mean.totalFat<-mean(USDAclean$TotalFat)
USDAclean$HighFat<-ifelse(USDAclean$TotalFat > mean.totalFat, 1, 0)
sum(USDAclean$HighSodium&USDAclean$HighFat)
```

#### 12. Calculate the average amount of iron, for high and low protein foods. (8 points)
```{r}
mean(USDAclean$Iron[USDAclean$HighProtein==1])
mean(USDAclean$Iron[USDAclean$HighProtein==0])
```

#### 13. Create a script for a "HealthCheck" program to detect unhealthy foods. Use the algorithm flowchart below as a basis for this script. (8 points)
```{r}
require(jpeg)
img<-readJPEG("HealthCheck.jpg")
plot(1:4, ty = 'n', ann = F, xaxt = 'n', yaxt = 'n')
rasterImage(img,1,1,4,4)
```
```{r}
HealthCheck <- function(food) {
  if(food[1]) {
    if(food[2]) {
      if(food[3]) {
        return("Fail")
      }
    }
  }
  return("Pass")
}
```

#### 14. Add a new variable called HealthCheck to the data frame using the output of the function. (8 points)
```{r}
#nrow(USDAclean[USDAclean$HighSodium&USDAclean$HighSugar&USDAclean$HighFat,])
USDAclean$HealthCheck<-apply(USDAclean[,c("HighSodium","HighSugar","HighFat")], 1, HealthCheck)
```

#### 15. How many foods in the USDAclean data frame fail the HealthCheck? (8 points)
```{r}
sum(USDAclean$HealthCheck=="Fail")
```

#### 16. Save your final data frame as "USDAclean_ [your last name]." (3 points)
```{r}
write.csv(USDAclean, "USDAclean_Zhang.csv")
```


This is the end of Assignment 1

Ceni Babaoglu, PhD
