---
title: 'CMTH 642 Data Analytics: Advanced Methods'
output:
  pdf_document: default
  html_document: default
  word_document: default
---
<center> <h1> Assignment 2 (10%) </h1> </center>
<center>  <h3> Rui Zhang </h2> </center>
<center> <h3> 500736315 </h2> </center>
---

#### 1. Read the csv file (USDA_Clean.csv) in the folder and assign it to a data frame. (3 points)
```{r}
# INSERT YOUR CODE HERE.
```

#### 2. Check the datatypes of the attributes. (3 points)
```{r}
# INSERT YOUR CODE HERE.
```

#### 3. Visualize the correlation among Calories, Protein, Total Fat, Carbohydrate, Sodium and Cholesterol. (7 points)
```{r}
# INSERT YOUR CODE HERE.
```

#### 4. Is the correlation between Calories and Total Fat statistically significant? Why? (7 points)
```{r}
# INSERT YOUR CODE HERE AND EXPLAIN YOUR ANSWER AS A COMMENT.
```

#### 5. Create a Linear Regression Model, using Calories as the dependent variable Protein, Total Fat, Carbohydrate, Sodium and Cholesterol as the independent variables. (7 points)
```{r}
# INSERT YOUR CODE HERE.
```

#### 6. Write the Linear Regression Equation, using Calories as the dependent variable whereas Protein, TotalFat, Carbohydrate, Sodium and Cholesterol as the independent variables. (7 points)
```{r}
# WRITE YOUR ANSWER AS A COMMENT.
```

#### 7. Which independent variable is the least significant? Why? (7 points)
```{r}
# INSERT YOUR CODE HERE AND EXPLAIN YOUR ANSWER AS A COMMENT.
```

#### 8. A new product is just produced with the following data:                                                            Protein=0.1, TotalFat=35, Carbohydrate=405, Sodium=440, Cholesterol=70, Sugar=NA, Calcium=35, Iron=NA, Potassium=35, VitaminC=10,	VitaminE=NA, VitaminD=NA.                                                                                                         Based on the model you created, what is the predicted value for Calories? (7 points)
```{r}
# INSERT YOUR CODE HERE.
```

#### 9. If the Sodium amount increases from 440 to 44440 (10000% increase), how much change will occur on Calories in percent? Explain why? (7 points)
```{r}
# INSERT YOUR CODE HERE AND EXPLAIN YOUR ANSWER AS A COMMENT.
```


#### 10. A study of primary education asked elementaty school students to retell two book articles that they read earlier in the week. The first (Article 1) had no pictures, and the second (Article 2) was illustrated with pictures. An expert listened to recordings of the students retelling each article and assigned a score for certain uses of language. Higher scores are better. Here are the data for five readers in this study:

#### Article 1   0.40   0.72   0.00   0.36   0.55

#### Article 2   0.77   0.49   0.66   0.28   0.38


#### A) What are $H_0$ and $H_a$ ? (5 points)
```{r}
# INSERT YOUR ANSWER AS A COMMENT.
```

#### B) Is this a paired or unpaired experiment? (5 points)
```{r}
# INSERT YOUR ANSWER AS A COMMENT.
```

#### C) Based on your previous answer, which nonparametric test statistic would you use to compare the medians of Article 1 and Article 2. (5 points) 
```{r}
# INSERT YOUR ANSWER AS A COMMENT.
```

#### D) Use a nonparametric test statistic to check if there is a statistically significant difference between the medians of Article 1 and Article 2. (5 points) 
```{r}
# INSERT YOUR CODE HERE.
```

#### E) Will you accept or reject your Null Hypothesis? ($\alpha = 0.05$) Do illustrations improve how the students retell an article or not? Why? (5 points)
```{r}
# INSERT YOUR ANSWER AS A COMMENT.
```

#### 11. Two companies selling toothpastes with the lable of 100 grams per tube on the package. We randomly bought eight toothpastes from each company A and B from random stores. Afterwards, we scaled them using high precision scale. Our measurements are recorded as follows:

#### Company A: 97.1   101.3   107.8   101.9   97.4    104.5   99.5    95.1
#### Company B: 103.5  105.3   106.5   107.9   102.1   105.6   109.8   97.2


#### A) Is this a paired or unpaired experiment? (5 points)
```{r}
# INSERT YOUR ANSWER AS A COMMENT.
```

#### B) Based on your previous answer, which nonparametric test statistic would you use to compare the medians of Company A and Company B. (5 points) 
```{r}
# INSERT YOUR ANSWER AS A COMMENT.
```

#### C) Use a nonparametric test statistic to check if there is a statistically significant difference between the medians of Company A and Company B. (5 points) 
```{r}
# INSERT YOUR CODE HERE.
```

#### D) Will you accept or reject your Null Hypothesis? ($\alpha = 0.05$) Are packaging process similar or different based on weight measurements? Why? (5 points)
```{r}
# INSERT YOUR ANSWER AS A COMMENT.
```

This is the end of Assignment 2

Ceni Babaoglu, PhD
